---
title: "Introduction to the simITS package"
author: "Luke Miratrix"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro simITS}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library( tidyverse )
library( simITS )
```

The raw Mecklenberg data we will use to illustrate the simITS package.
```{r}
data(mecklenberg)
head( mecklenberg )
meck = mutate( mecklenberg, pbail = 100 * pbail )
ggplot( meck, aes( x=month, y=pbail)) +
  geom_rect(aes( ymin=-Inf, ymax=Inf, xmin=0.5, xmax=25, fill="lightgray"), col = "lightgray", alpha=0.25) +
  scale_fill_identity(name = "", guide = "none", labels = c('Post Policy era')) +
  geom_hline( yintercept = 0, col="black") +
  geom_line( col="black", lty=1, lwd=0.5) +
  geom_point() +
  scale_x_continuous( breaks = c(-29,-24,-18,-12,-6,1,6,12,18,24)) +    
  coord_cartesian(xlim=c(-29.5,24.5), ylim=c(0,100), expand=FALSE) +
  labs( title = " ", y = "Percent cases assigned bail", x = " " )
```

We can add lagged covariates and fit our simple ITS model as so:
```{r}
    meck = add_lagged_covariates( meck, outcomename = "pbail", covariates=NULL )
    head( meck )
    tail( meck )
    
    meck.pre = filter( meck, month <= 0 )    
    mod = fit_model.default( meck.pre, "pbail" )
    summary( mod )
```

```{r}
  t0 = 0
    envelope = process_outcome_model( "pbail", meck, 
                                      t0=t0, R = 100, 
                                      summarize = TRUE, smooth=FALSE )
    head( envelope )
    tail( envelope )

    ggplot( envelope, aes( month ) ) +
        geom_line( aes( y=Y ), alpha = 0.6 ) +  # original data
        geom_point( aes( y=Y ) ) +              # original data
        #    geom_line( aes( y=Ystar1 ), alpha=0.7, col="green" ) +   # loess smoothed line
        geom_ribbon( aes( ymin=Ymin, ymax=Ymax ), alpha=0.2 ) +
        geom_line( aes( y = Ystar ), col="darkgrey" ) +
        geom_vline( xintercept = t0+0.5)
```


# Demo of smoothing and summarization ####
 
For simulation we also offer a fake data generator.  It works like this:
```{r}
  library( simITS )

    dat = make_fake_data( t_min=-60, t_max=18, t0 = 0 )
    head( dat )
    qplot( month, Ystr, data=dat, geom = c( "point","line") )
    qplot( month, Y, data=dat, geom = c( "point","line") )


```

We can then process it.  Here we demonstrate summarizing and smoothing:   
```{r}

    #debugonce( process_outcome_model )
    envelope = process_outcome_model( "Y", dat, t0=t0, R = 100, summarize = TRUE, smooth=TRUE )
    head( envelope )
    tail( envelope )

    make_envelope_graph(envelope, t0 )

    ggplot( dat, aes( month ) ) +
        geom_line( aes( y=Y ), alpha = 0.6 ) +
        geom_point( aes( y=Y ) ) +
        geom_line( data=envelope, aes( y=Ysmooth1 ), alpha=0.7, col="green" ) +
        geom_ribbon( data=envelope, aes( ymin=Ymin, ymax=Ymax ), alpha=0.2 ) +
        geom_line( data=envelope, aes( y = Ysmooth ), col="darkgrey" ) +
        geom_vline( xintercept = t0+0.5)


```

# Seasonality and covariates

A seasonality model on some fake data with a strong seasonality component is easy to fit.  You just construct some code to fit the seasonality model via the `make_fit_season_model()` factory (you need to have the covariates pre-constructed in your data):
```{r}
data( newjersey )
head( newjersey )
fit.season.model.qtemp =  make_fit_season_model( ~ temperature + Q2 + Q3 + Q4 )

envelope = process_outcome_model( "n.warrant", newjersey, t0=-7, R = 100, 
                                  summarize = TRUE, smooth=TRUE, 
                                  fit_model = fit.season.model.qtemp )
head( envelope )
tail( envelope )

make_envelope_graph( envelope, t0=-7 )
```

Note how it will construct the lagged covariates automatically. The `make_fit_season_model()` method records what covariates are needed from the passed formula.



